Indicators for Ecosystem Condition in Norway

### 2023-06-15 DRAFT_ndvi_wetland.Rmd
sdlfksdf

#### NiN data
We read the NiN data for this document from a cache at the NINA server...
```{r, echo = F}
# Add data from cache
nin<-readRDS(paste0(here::here(),"/data/cache/nin.RDS"))
```

...but it can also be directly downloaded from Miiljødirektoratets kartkatalog (https://kartkatalog.miljodirektoratet.no/Dataset)
```{r, echo = T, eval = F}

# st_layers(dsn = "/data/P-Prosjekter2/41201785_okologisk_tilstand_2022_2023/data/Naturovervaking_eksport.gdb")
# ANO.sp <- st_read("/data/P-Prosjekter2/41201785_okologisk_tilstand_2022_2023/data/Naturovervaking_eksport.gdb",
#                    layer="ANO_Art")
# ANO.geo <- st_read("/data/P-Prosjekter2/41201785_okologisk_tilstand_2022_2023/data/Naturovervaking_eksport.gdb",
#                   layer="ANO_SurveyPoint")
# head(ANO.sp)
# head(ANO.geo)
```

Plant indicators from Tyler et al. (2021)
```{r, echo = F}
## Tyler indicator data
## data from cache
ind.dat <-readRDS(paste0(here::here(),"/data/cache/ind.dat.RDS"))
# ind.dat <- read.table("/data/P-Prosjekter2/41201785_okologisk_tilstand_2022_2023/data/functional plant indicators/Tyler et al_Swedish plant indicators.txt",
#                         sep = '\t', header=T, quote = '')
# head(ind.dat)

>>>>

#### NiN data
We read the NiN data for this document from a cache at the NINA server...
```{r, echo = F}
# Add data from cache
nin <- readRDS(paste0(here::here(),"/data/cache/nin.RDS"))
```

...but it can also be directly downloaded from Miljødirektoratets kartkatalog (https://kartkatalog.miljodirektoratet.no/Dataset)
```{r, echo = T, eval = F}

url <- "https://nedlasting.miljodirektoratet.no/Miljodata//Naturtyper_nin/FILEGDB/4326/Naturtyper_nin_0000_norge_4326_FILEGDB.zip"
download(url, dest="P:/41201785_okologisk_tilstand_2022_2023/data/Naturtyper_nin_0000_norge_4326_FILEGDB.zip", mode="wb") 
unzip ("P:/41201785_okologisk_tilstand_2022_2023/data/Naturtyper_nin_0000_norge_4326_FILEGDB.zip", 
       exdir = "P:/41201785_okologisk_tilstand_2022_2023/data/nin_data")
st_layers("P:/41201785_okologisk_tilstand_2022_2023/data/nin_data/Naturtyper_nin_0000_norge_4326_FILEGDB.gdb")
nin <- st_read("P:/41201785_okologisk_tilstand_2022_2023/data/nin_data/Naturtyper_nin_0000_norge_4326_FILEGDB.gdb",
                 layer="naturtyper_nin_omr")

#### Geometry for Norway and the five regions

```{r}
# Import region- og Norgeskart, make sure they have the same geometry as the nin data
nor <- st_read("data/outlineOfNorway_EPSG25833.shp")%>%
  st_as_sf() %>%
  st_transform(crs = st_crs(nin))

reg <- st_read("data/regions.shp")%>%
  st_as_sf() %>%
  st_transform(crs = st_crs(nin))

# combine the Norway and the region maps
regnor <- st_intersection(reg,nor)
```

#### NDVI data
Plant indicators from Tyler et al. (2021)
```{r, echo = F}
## Tyler indicator data
## data from cache
ind.dat <-readRDS(paste0(here::here(),"/data/cache/ind.dat.RDS"))
# ind.dat <- read.table("/data/P-Prosjekter2/41201785_okologisk_tilstand_2022_2023/data/functional plant indicators/Tyler et al_Swedish plant indicators.txt",
#                         sep = '\t', header=T, quote = '')
# head(ind.dat)
